---
title: "Time Traveler (Group 5) Final Project - Demand Forecasting Using Seasonal ARIMA: Model Fit and Forecast Accuracy."  
author: "Youngjae Cho"
date: "2025-04-23 ~ 2025-05-01, 2025-05-05 ~ 2025-05-07, 2025-05-08 ~ 2025-05-10"
output: 
  html_document:
    self_contained: true
---

```{r setup, include=FALSE, purl=FALSE}
# Use echo = FALSE for Portfolio assignments
knitr::opts_chunk$set(echo = FALSE)
```

```{r metadata, echo=FALSE}
# Author:  Youngjae Cho
# Date:    2025-04-23 ~ 2025-05-01, 2025-05-05 ~ 2025-05-07, 2025-05-08 ~ 2025-05-10
# Purpose: R Markdown file for Final Project (STAT 4720)
#---------------------------------------------------------------------
```


```{r packages}
# Load packages
suppressPackageStartupMessages(library("tidyverse")); theme_set(theme_bw())
library("knitr")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Library setting
library(RTseries)
```

## Purpose: 
In Milestone 1, at the last printed R result, we have seen the seasonality. 
In this Milestone 2, we used a range-mean plot to check the transformation and iden() function to 
check the stationarity for 4 differencing schemes. 
First off, I will load the data set and build the original time series plot that used in Milestone.

## Data 

```{r}
miles2 <- read.csv("Daily Demand Forecasting Orders.csv") # Data source. 
miles2 

# For seeing the first few rows to check.
head(miles2)

# To check the column for variables.
colnames(miles2)

# To generate a sequential time index
miles2$time_index <- seq(from=1, to = nrow(miles2), by =1)
miles2$time_index

# Select the total orders variables.
miles2_ts <- ts(miles2$Target..Total.orders., frequency = 5) # Weekly frequency for 5 business days.
miles2_ts # Start from 1 to 60. 

# Time Series Plot.
plot(miles2_ts, type="l", col="salmon", lwd=2,
     main = "Time Series Plot - Realization vs Time",
     xlab = "Day", ylab = "Total Orders")
```

```{r miles2}
## In this Milestone2, the time series plot between Realization and Time, we can see the seasonality.
## Therefore, I will use s = 5 for five business days, that is weekly cycles, Monday to Friday. 
## I will use acf, pacf plots after log-transformation. 

# Base functions for ACF & PACF comparison
par(mfrow=c(2,2))

# ACF & PACF plots for original time series data.
acf(miles2_ts, main="ACF: Original Data")
pacf(miles2_ts, main="PACF: Original Data")

# Log-transformed data
log_miles2_ts <- log(miles2_ts)

# ACF & PACF plots of log-transformed data
acf(log(miles2_ts), main="ACF: Log-Transformed Data")
pacf(log(miles2_ts), main="PACF: Log-Transformed Data") # No need for transformation. 

par(mfrow=c(1,1)) 
```
```{r}
## Then, I will use iden() function to check the stationary. 
# Run the Seasonal identification with two different differencing schemes.
miles2_tsd <- tsd(miles2_ts,
                  data.title = "Daily Demand Forecasting Orders",
                  response.units = "Orders",
                  time.units = "Day")
miles2_tsd

# iden function for only two differencing in seasonal time series. 
iden(miles2_tsd, d = 0, D = 1) # Seasonal only.
iden(miles2_tsd, d = 1, D = 1) # First order and seasonal but it would be over-differencing. 

```

#4. Library setting
```{r library setting}
library(forecast)
library(TSA)
library(tseries)
```

#5. Do ADF Test for getting Stationary and fit four candidate SARIMA Models. (5/5)

# (0, 0, 2)(1, 0, 1)_5, residual analysis, forecast are needed for further steps. 

# Interpretation: 
After adf test, I got p-value of 0.01 from the (d=1, D=1) differenced series. The p-value was smaller than the signficance level so we can reject the null hypothesis. In conclusion, the series is stationary after conducting the first and seasonal differencing. 
For this part: Fit 2 has larger AIC, BIC than other fit models. Therefore, I decide fit2 for model selection. 

```{r}
# data 
data <- read.csv("Daily Demand Forecasting Orders.csv")
ts_data <- ts(data$Target..Total.orders., frequency = 5) # For weekly seasonality.

print(data)
print(ts_data)

# ADF Tests for two differenced series
cat("ADF Test for Seasonal Differencing Only (d=0, D=1):\n") # No - Non-Stationary. (High p-value)
adf.test(diff(ts_data, lag = 5), alternative = "stationary")

cat("\nADF Test for First and Seasonal Differencing (d=1, D=1):\n") # Yes - Stationary Confirm! (small p-value)
adf.test(diff(diff(ts_data), lag = 5), alternative = "stationary")

# Fit four candidate SARIMA Models. 
fit1 <- Arima(ts_data, order =c(0,1,1), seasonal= list(order =c(0,0,1), period= 5))
fit2 <- Arima(ts_data, order =c(1,1,1), seasonal= list(order =c(0,1,1), period= 5))
fit3 <- Arima(ts_data, order =c(0,1,2), seasonal= list(order =c(1,0,1), period = 5))
fit4 <- Arima(ts_data, order =c(0,0,2), seasonal= list(order =c(1,0,1), period= 5))

print(fit1)
print(fit2)
print(fit3)
print(fit4)

```

#6. Compare AIC and BIC. 

# Interpretation for 6.
Why the model 2 is the best fit? Why 4? 
First off, Model 2 AIC, BIC values are smaller than other models. 
Therefore, the model 2 is the best fit. 

```{r}
cat("Model 1 (0,1,1)(0,0,1)[5]:\n", "AIC:",AIC(fit1), "BIC:", BIC(fit1),"\n")
cat("Model 2 (1,1,1)(0,1,1)[5]:\n", "AIC:",AIC(fit2), "BIC:", BIC(fit2),"\n")
cat("Model 3 (0,1,2)(1,0,1)[5]:\n", "AIC:",AIC(fit3), "BIC:", BIC(fit3),"\n")
cat("Model 4 (0,0,2)(1,0,1)[5]:\n", "AIC:",AIC(fit4), "BIC:", BIC(fit4),"\n")
```

#7. Residual diagnostics for fit4 (SARIMA(0,0,2)(1,0,1)[5]) AND Fit2 (SARIMA(1,1,1)(0,1,1)[5])

# Interpretation: 
I think Model 2(fit 2) is better to fit because the AIC, BIC values from fit2 are smaller than other 3 models. 
Reasons: 
- lower AIC, BIC values.
- P-value was larger than significance level due to p-value = 0.2474. 
- Better generation for forecasting. 
- Most data was close to red line, so reasonable Q-Q normality. (But there were some outliers) 
- Proper differencing (d=1, D=1) for dealing with Stationarity. 
Conclusion: Model 2 is well-fitted and appropriate for SARIMA Model for this data set. 


Model 4 has the highest AIC and BIC values and in the Normal Q-Q plot, it looks more skewed. 
```{r}
# Model 4, SARIMA(0,0,2)(1,0,1)[5] model for Residual Diagnostics.
# par(mfrow = c(2, 1))
# acf(residuals(fit4), main = "ACF of Residuals")
# pacf(residuals(fit4), main = "PACF of Residuals")
# par(mfrow = c(1, 1))

# Box.test(residuals(fit4), lag = 10, type = "Ljung-Box", fitdf = 3)  # df = 3.
# qqnorm(residuals(fit4)); qqline(residuals(fit4), col = "blue")

# Best Model: Model 2, SARIMA(1,1,1)(0,1,1)[5] model for Residual Diagnostics.
par(mfrow = c(2, 1))
acf(residuals(fit2), main = "ACF of Residuals(fit2)")
pacf(residuals(fit2), main = "PACF of Residuals(fit2)")
par(mfrow = c(1, 1))

Box.test(residuals(fit2), lag = 10, type = "Ljung-Box", fitdf = 3)
qqnorm(residuals(fit2)); qqline(residuals(fit2), col = "red")
```
#8. Forecasting with Model 2 (SARIMA(1,1,1)(0,1,1)[5])
#Interpretation: 
This model 2 maintains the weekly cycle in the original data with period = 5. There were no big growth or decline. And this indicates that there would be stable seasonality and trend behavior. 
In short term forecasts, the prediction intervals are relatively narrow. It indicates the high confidence in the short-term forecast. The model projects the same seasonal up and down pattern. 
Also, in the medium-term forecasts, for example, 1 month to 5 month, the seasonal cycle is preserved. Uncertainty bands widen moderately for further ahead. At the long-term forecast, for instance, 1 year, the forecast still preserved the seasonal pattern but with much wider confidence intervals. I think this shows the realistic modeling. There were no strong unrealistic drift so the model 2 would be the most suitable SARIMA model to forecast. 
As a result, SARIMA(1,1,1)(0,1,1)[5] makes a stable and seasonally appropriate forecasting. 
The model is consistent with previous data and contains the reasonable uncertainty bands. 
Therefore, the model 2 is the most suitable choice for forecasting. 

```{r}
#1. Forecast for the next 5 business days. (Short-term)
forecast_result_fit2 <- forecast(fit2, h = 5, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 5 days ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#2. Forecast for the next 10 days. 
forecast_result_fit2 <- forecast(fit2, h = 10, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 10 days ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#3. Forecast for the next 20 days.
forecast_result_fit2 <- forecast(fit2, h = 20, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 20 days ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#4. Forecast for the next 30 days. (1 month)
forecast_result_fit2 <- forecast(fit2, h = 30, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 1 month ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#5. Forecast for the next 60 days. (2 month)
forecast_result_fit2 <- forecast(fit2, h = 60, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 2 month ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#6. Forecast for the next 90 days. (3 month)
forecast_result_fit2 <- forecast(fit2, h = 90, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 3 month ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#7. Forecast for the next 120 days. (4 month)
forecast_result_fit2 <- forecast(fit2, h = 120, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 4 month ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#8. Forecast for the next 150 days. (5 month)
forecast_result_fit2 <- forecast(fit2, h = 150, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 5 month ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#9. Forecast for the next 180 business days. (6 months)
forecast_result_fit2 <- forecast(fit2, h = 180, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 6 month ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

#10. Forecast for the next 365 days. (1 year)
forecast_result_fit2 <- forecast(fit2, h = 365, level = c(80, 95))
plot(forecast_result_fit2, main = "Forecast: SARIMA(1,1,1)(0,1,1)[5] - 1 year ahead", xlab = "Day", ylab = "Forecasted Orders")
print(forecast_result_fit2)

```
#9. External Variables for improving process.

# Interpretation: 

I wanted to look the difference between the original seasonal time series and time series with external variables. 
The regression SARIMA Model 2 forecast for the next 60 days with urgent orders and non-urgent orders has a similar visualization with original SARIMA Model. 
I think the original SARIMA Model is strong and well-indicated. 
The SARIMA Model 2 has lower AIC, BIC values but they are really similar visualization. Therefore, it did not change the main seasonal pattern and forecast trend. 
In conclusion, the primary seasonal ARIMA model explains the most of the variation in total orders and the external variables did not do well but the statistical goodness of fitness measurement was improved based on lower AIC, BIC values. 

```{r}
# Data set setting.
ts_data <- ts(miles2$Target..Total.orders., frequency = 5) # Original data.

# Add two external objects, urgent and non-urgent orders.
nonurgent_ts <- miles2$Non.urgent.order
urgent_ts <- miles2$Urgent.order
xreg_vars <- cbind(urgent_ts, nonurgent_ts) # Design matrix for two external variables. 

print(nonurgent_ts)
print(urgent_ts)
print(xreg_vars)

# Fit the SARIMA Model2 with external variables - dynamic regression.
fit_regression_sarima2 <- Arima(ts_data,
                                order = c(1,1,1),
                                seasonal = list(order = c(0,1,1), period = 5),
                                xreg = xreg_vars)
fit_regression_sarima2

summary(fit_regression_sarima2) # lower AIC, BIC than original SARIMA Model.

# Check process for residual diagnostics.
par(mfrow = c(2,1))
acf(residuals(fit_regression_sarima2), main = "ACF of Residuals (Regression SARIMA 2)") # White Noise again.
pacf(residuals(fit_regression_sarima2), main = "PACF of Residuals (Regression SARIMA 2)") # Similar with ACF - For two plots, there is no meaningful autocorrelation in 
                                                                                          # the residuals.
par(mfrow = c(1,1))

Box.test(residuals(fit_regression_sarima2), lag = 10, type = "Ljung-Box") # Higher P-value than significance level - good.

# Normal Q-Q Plot.
qqnorm(residuals(fit_regression_sarima2)); qqline(residuals(fit_regression_sarima2), col = "black")

# Forecast with external variables for the next 60 days.
future_nonurgent <- rep(mean(nonurgent_ts), 60)
future_urgent <- rep(mean(urgent_ts), 60)
future_xreg <- cbind(future_urgent, future_nonurgent)

forecast_regression_sarima2 <- forecast(fit_regression_sarima2,
                                        h = 60,
                                        xreg = future_xreg,
                                        level = c(80,95))
# Visualization.
plot(forecast_regression_sarima2, main = "Regression SARIMA Model2 Forecast for the next 60 Days",
     xlab = "Day", ylab = "Forecasted Total Orders")
```
#10. Report for Seasonal ARIMA Modeling with "Daily Demand Forecasting Orders" data set: 

Conclusion: The SARIMA Model 2 with original data is the most appropriate suitable model for forecasting seasonal demand due to statitsically appropriate fit. This is based on Model 2 having the lowest AIC, BIC, p-value, proper stationarity handling in (d=1, D=1), and reasonable normality. The result between regression SARIMA Model2 with external variables and SARIMA Model 2 with original data is that the SARIMA Model 2 with original data already shows the most of seasonal behavior and trend in total orders. Despite the SARIMA Model 2 with external variables has a lower AIC, BIC than the model with original data, the visualizations between them are visually similar. This indicates that adding two external variables is not helpful for changing the forecast trajectory. However, this is helpful for improving the statistical goodness of fitness. 
There are some outliers in Normal Q-Q Plot and uncertainty increases in forecasting from the SARIMA Model 2 with original data. In short-term, this model is the most suitable but in long-term, using the model 2's forecasting causes less precise after next 1 year ahead because confidence intervals gradually widen over time. 
Overall, our project concludes that the Model 2, SARIMA(1,1,1)(0,1,1)[5] is the most appropriate model for forecasting seasonal demand. But, if we have more data with external variables, such as, promotions and holidays, it might be helped in precision for a long-term. 
Thank you!









