---
title: "Portfolio 7 - Warpbreaks Analysis"
author: "Youngjae Cho"
date: "2024-02-08 ~ 2024-04-01"
output: 
  html_document:
    self_contained: true
---

```{r setup, include=FALSE, purl=FALSE}
# Use echo = FALSE for Portfolio assignments
knitr::opts_chunk$set(echo = FALSE)
```

```{r metadata, echo=FALSE}
# Author:  Youngjae Cho
# Date:    2024-02-08 ~ 2024-04-01
# Purpose: R Markdown file with P7 - Warpbreaks Analysis 
#-------------------------------------------------------------------------------
```


```{r packages}
# Load packages
suppressPackageStartupMessages(library("tidyverse")); theme_set(theme_bw())
library("knitr")
```

The purpose of this document is as an template R Markdown file that can be used
as an example for constructing portfolio assignments in this course. 
For demonstration purposes, we will utilize the `warpbreaks` data set. 

## Introduction

The warpbreaks data set consists of `r nrow(warpbreaks)` observations from a 
designed experiment. 
This experiment aims to study how wool type and tension affect the number
of breaks observed during weaving. 
In particular, we are interested in the wool type and tension that minimizes
the number of breaks. 
Since tension is coded as categorical variable (as opposed to numeric), 
we will be limited to these three choices (L/M/H) for tension in minimizing
the number of breaks.

## Data

```{r summary-statistics}
# Calculate summary statistics by wool-tension combination
s <- warpbreaks |>
  group_by(wool, tension) |>
  summarize(
    n = n(),
    mean = mean(breaks),
    sd   = sd(breaks),
    
    .groups = "drop"
  ) |>
  mutate(
    mean_sd = paste0(
      format(round(mean, 1), nsmall = 1), # Keep exactly 1 decimal place
      " (",
      format(round(sd,   1), nsmall = 1), # Keep exactly 1 decimal place
      ")"
    )
  ) 
```

The experiment was *balanced* with `r unique(s$n)` replicates for each 
combination of wool type (A/B) and tension (L/M/H). 
The number of breaks ranged from `r min(warpbreaks$breaks)` to 
`r max(warpbreaks$breaks)` across all replicates. 
The table below provides the mean and standard deviation of the number of 
breaks for each combination of wool and tension. 
As can be seen in this table, 
wool B and tension H produces the fewest breaks and the least variability
in the number of breaks.


```{r summary-statistics-table}
# Create HTML table from summary statistics
s |>
  pivot_wider(
    id_cols = wool,
    names_from = tension,
    values_from = mean_sd
  ) |>
  rename(
    Wool = wool
  ) |>
  knitr::kable(
    caption = "Mean (SD) for Number of Breaks for each combination of Wool and Tension",
    # label   = "summary-statistics",
    align   = c("rlll")) # Column alignment: [r]ight or [l]eft
```

The figure below displays a scatterplot of the number of breaks versus tension
and wool type. 
When comparing wool A to wool B, 
this figure suggests wool A has more breaks with tension L, 
about the same mean breaks for tension M, 
and (perhaps) fewer mean breaks for tension H. 
For wool B and tension H, number of breaks ranged from 
`r warpbreaks |> filter(wool=="B", tension=="H") |> pull(breaks) |> min()` to
`r warpbreaks |> filter(wool=="B", tension=="H") |> pull(breaks) |> max()`.

```{r exploratory}
# Create exploratory plot of the data
ggplot(warpbreaks,
       aes(
         x = tension,
         y = breaks,
         color = wool,
         shape = wool
       )) +
  geom_point(
    position = position_jitterdodge(
      dodge.width  = 0.1,
      jitter.width = 0.05
    )
  ) +
  labs(
    x = "Tension",
    y = "Number of Breaks",
    title = "Yarn Weaving Experiment"
  )
```

## Results

```{r model-interaction}
# Fit Poisson regression model with interaction
m <- glm(breaks ~ wool * tension,
         data = warpbreaks,
         family = "poisson")
```

Due to number of breaks being a count, 
we will fit a Poisson regression model for number of breaks on
wool, tension, and their interaction.
The Poisson regression model assumes 
$$Y_i \stackrel{ind}{\sim} Po(\lambda_i)$$
where $Y_i$ is the number of breaks. The Poisson rate parameter is related to
the wool type and tension through 
$$\log(\lambda_i) = \beta_0 + \beta_1 \mbox{woolB}_i +
\beta_2 \mbox{tensionM}_i + \beta_3 \mbox{tensionH}_i +
\beta_4 \mbox{woolB}_i\times \mbox{tensionM}_i +
\beta_5 \mbox{woolB}_i\times \mbox{tensionH}_i$$
where, for observation $i$, 
$\mbox{woolB}_i$ is an indicator for wool type B, 
$\mbox{tensionM}_i$ is an indicator for tension M, and 
$\mbox{tensionH}_i$ is an indicator for tension H.

Since the interaction between wool and tension is significant 
(p $\approx$ `r format(round(anova(m, test="Chi")[4,5], 4), nsmall=4)`), 
we will keep the interaction in the model. 

A plot of the estimated mean number of breaks and 95% confidence
intervals is provided below. 
As we saw with the original data, 
the number of breaks generally decreases as tension increases.
For a given tension, there are typically fewer breaks with 
wool B compared to wool A. 

```{r predictions}
# Create data frame with unique explanatory variable values
nd <- warpbreaks |>
  select(tension, wool) |>
  unique()

# Predict
p <- bind_cols(
  nd,
  predict(m,
          newdata = nd,
          se.fit = TRUE)|>
      as.data.frame() |>
      
      # Manually construct confidence intervals
      mutate(
        lwr = fit - qnorm(0.975) * se.fit,
        upr = fit + qnorm(0.975) * se.fit,
        
        # Exponentiate to get to response scale
        breaks = exp(fit),
        lwr    = exp(lwr),
        upr    = exp(upr)
      ) 
  )

# Plot group estimates
pg <- ggplot(p,
             aes(
               x = tension,
               y = breaks,
               color = wool,
               shape = wool,
               group = wool,
               ymin = lwr,
               ymax = upr
             )) + 
  geom_pointrange() +
  geom_line() +
  labs(
    x = "Tension",
    y = "Number of Breaks",
    title = "Estimated Mean Breaks with 95% Confidence Intervals",
    subtitle = "Poisson Regression Model"
  )

pg
```

## Discussion

```{r ci}
# Construct confidence interval to answer scientific question
ci <- p |>
  filter(wool == "B",
         tension == "H") |>
  select(lwr,upr) |>
  mutate(
    lwr = round(lwr, 1),
    upr = round(upr, 1)
  )

```

We investigated a designed experiment to assess the impact of the wool type and
tension on the number of breaks. 
We found that in these data wool type B and tension H is resulted in the fewest
mean breaks.
A 95\% confidence interval for the mean number of breaks with wool type B and 
tension H is (`r ci$lwr`, `r ci$upr`).

The data were likely obtained sequentially and thus there is the chance for a 
temporal effect violating our independence assumption, but no evidence of this
was found in our diagnostic plots. 
While the data are counts, and thus the normality assumption is technically 
violated, qq-plots of the residuals indicate the normality assumption is 
reasonable. 
Compared to the other five combinations of tension and wool, 
the low tension, wool A may have slightly more variability. 
As this group has only slightly more variability, we are not worried about the
constant variance assumption. 
Alternative approaches would be to use the logarithm of breaks as the response 
variable or use Poisson regression. 

From the information we have, it is unclear if any randomization of wool or 
tension to a particular run was done. 
If wool and tension were randomly assigned to a particular run,
then we can claim that the wool type and tension **caused** the difference in 
the number of breaks.
Even if randomization was done, our causal claims would be stronger for 
tension than wool since we can only randomly assign wool to a run and not the
type of wool to a bundle of wool.
Since this yarn and looms are not a random sample from any population, 
we need to be careful in trying to extend these results to some larger 
population.


